---
title: "Regression Models"
author: "Ajith Masthan"
date: "May 17, 2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

By looking at a data set of a collection of cars we are interested in exploring the relationship between a set of variables and miles per gallon. In particularly the following two questions:

* “Is an automatic or manual transmission better for MPG”
* "Quantify the MPG difference between automatic and manual transmissions"

The following steps are followed

- Data Processing
- Data Exploration of variables Transmission and MPG
- Model selection by trying out different models
- Model validation
- Final conclusions on results observed

## Data Processing

Changing am and cylinders to factors

```{r}
library(ggplot2)
library(GGally)
library(dplyr)
library(ggfortify)

data(mtcars)

mtcarsFactors <- mtcars
mtcarsFactors$am <- as.factor(mtcarsFactors$am)
levels(mtcarsFactors$am) <- c("automatic", "manual")

mtcarsFactors$cyl <- as.factor(mtcarsFactors$cyl)
mtcarsFactors$gear <- as.factor(mtcarsFactors$gear)
mtcarsFactors$vs <- as.factor(mtcarsFactors$vs)
levels(mtcarsFactors$vs) <- c("V", "S")
```

## Exploratory data analyses 

Basic Data summary
```{r}
# Res 1
dim(mtcarsFactors)

# Res 2
head(mtcarsFactors)
```

Relation between parameters of interest

```{r}
# Figure 1
library(ggplot2)
p <- ggplot(mtcarsFactors, aes(am, mpg))
p + geom_boxplot(aes(fill = am))
```

Observation: The manual transmissions have higher mpg's.  

Correlations of all the other variables are observed to include those with correlation higher than am

```{r}
# Res 3
cors <- cor(mtcars$mpg, mtcars)
orderedCors <- cors[,order(-abs(cors[1,]))]
orderedCors

# Res 4
amPos <- which(names(orderedCors)=="am")
subsetColumns <- names(orderedCors)[1:amPos]
subsetColumns

# Figure 2
mtcarsFactors[,subsetColumns] %>% 
   ggpairs(
     mapping = ggplot2::aes(color = am), 
     upper = list(continuous = wrap("cor", size = 3)),
     lower = list(continuous = wrap("smooth", alpha=0.4, size=1), combo = wrap("dot"))
   )
```


## Model selection 

Observation : Many variables have strong correlation other than am.

Need to include these variables to have an accurate model

Basic Model
```{r}
# Res 5
basicFit <- lm(mpg ~ am, mtcarsFactors)
summary(basicFit)
```

Observation: P Value is low, Rsquared is also not satisfactory

Including all the variables in the model
```{r}
# Res 6
totalFit <- lm(mpg ~ ., mtcarsFactors)
summary(totalFit)
```
Observation:Rsquared value is improved. P value doesn't show any significance

Trying StepWise Regression

```{r}
# Res 7
bestFit <- step(totalFit,direction="both",trace=FALSE)
summary(bestFit)
```

## Model examination

The best model resulting from Step wise Regression is `mpg ~ wt + qsec + am` 
P Values are significant for all the 3 variables
Rsquared is also high

Investigating Residuals Vs fitted

```{r}
# Figure 3
autoplot(bestFit)
```

Observation: 'Normal Q-Q' plot is good, but the 'Residuals vs Fitted' and 'Scale-Location' are not satisfactory

## Conclusion

Question : “Is an automatic or manual transmission better for MPG” 
All the models experimented explain manual transmission will increase your MPG holding all other paramters constant

Question : “Quantify the MPG difference between automatic and manual transmissions”
Manual transmission has 3 miles per gallon more than automatic.
This is based on the result from the best fit model `mpg ~ wt + qsec + am` with p value <0.05 and Rsquared = 0.85

Other Observation :
Model is not accurate as suggested by Residuals Vs. Fit plot
This ight be due to lower number of observations (32) in the data

Though it can be said Manual Transmissions gives 3 MPG more than Automatic. I might not be accurate for any of the future data.

